Constructing a universe for the setoid model

### 2020-07-08 paper.tex
ERC support + compression

\author{Thorsten Altenkirch}{School of Computer Science, University of Nottingham, UK}{psztxa@nottingham.ac.uk}{https://orcid.org/0000-0002-6582-5025}{Supported by USAF grant FA9550-16-1-0029.}
\author{Simon Boulier}{Inria, Nantes, France}{simon.boulier@inria.fr}{}{}
\author{Ambrus Kaposi}{E{\"o}tv{\"o}s Lor{\'a}nd University, Budapest, Hungary}{akaposi@inf.elte.hu}{https://orcid.org/0000-0001-9897-8936}{Supported by the National Research, Development and Innovation Fund of Hungary, financed under the Thematic Excellence Programme funding scheme, Project no. ED\_18-1-2019-0030 and the Bolyai Fellowship of the Hungarian Academy of Sciences, Project no.\ BO/00659/19/3.}
\author{Christian Sattler}{Chalmers University of Technology, Gothenburg, Sweden}{sattler@chalmers.se}{}{Supported by USAF grant FA9550-16-1-0029 and Swedish Research Council grant 2019-03765.}
\author{Filippo Sestini}{School of Computer Science, University of Nottingham, UK}{Filippo.Sestini@nottingham.ac.uk}{}{}

@@

We have described the construction of a universe of setoids in the setoid model
of type theory; this is given in several steps, first as an inductive-recursive
definition, then as an inductive-inductive definition, and finally as an
inductive type. Every encoding is obtained from the previous by adapting known
data type transformation methods in a novel way that accounts for the
peculiarities of our construction.

Most of the mathematical contents of this paper have been formalized in the
proof-assistant Agda (see \cite{agda-code}). We aim to complete the
formalization in the near future, in particular to obtain a machine-checked
proof that our inductive construction of the setoid universe does not rely on
recursion-recursion.

In contrast to the inductive-recursive and inductive-inductive
versions of the universe, the inductive definition relies on a
metatheory with a strong transport rule. As future work, we would like
to prove normalization for this metatheory, since we don't yet know if
this property holds in the presence of such rule, and previous work in
this respect \cite{abel2019failure} seems to suggest that is
represents a non-trivial addition.

>>>>

\author{Thorsten Altenkirch}{School of Computer Science, University of Nottingham, UK}{psztxa@nottingham.ac.uk}{https://orcid.org/0000-0002-6582-5025}{Supported by USAF grant FA9550-16-1-0029.}
\author{Simon Boulier}{Inria, Nantes, France}{simon.boulier@inria.fr}{}{Supported by ERC Starting Grant CoqHoTT 637339.}
\author{Ambrus Kaposi}{E{\"o}tv{\"o}s Lor{\'a}nd University, Budapest, Hungary}{akaposi@inf.elte.hu}{https://orcid.org/0000-0001-9897-8936}{Supported by the National Research, Development and Innovation Fund of Hungary, financed under the Thematic Excellence Programme funding scheme, Project no. ED\_18-1-2019-0030 and the Bolyai Fellowship of the Hungarian Academy of Sciences, Project no.\ BO/00659/19/3.}
\author{Christian Sattler}{Chalmers University of Technology, Gothenburg, Sweden}{sattler@chalmers.se}{}{Supported by USAF grant FA9550-16-1-0029 and Swedish Research Council grant 2019-03765.}
\author{Filippo Sestini}{School of Computer Science, University of Nottingham, UK}{Filippo.Sestini@nottingham.ac.uk}{}{}

@@

We have described the construction of a universe of setoids in the setoid model
of type theory; this is given in several steps, first as an inductive-recursive
definition, then as an inductive-inductive definition, and finally as an
inductive type. Every encoding is obtained from the previous by adapting known
data type transformation methods in a novel way that accounts for the
peculiarities of our construction.
%
Most of the mathematical contents of this paper have been formalized in the
proof-assistant Agda (see \cite{agda-code}). We aim to complete the
formalization in the near future, in particular to obtain a machine-checked
proof that our inductive construction of the setoid universe does not rely on
recursion-recursion.

In contrast to the inductive-recursive and inductive-inductive
versions of the universe, the inductive definition relies on a
metatheory with a strong transport rule. As future work, we would like
to prove normalization for this metatheory since previous work in
this respect \cite{abel2019failure} seems to suggest that is
represents a non-trivial addition.
